Arne report

### 2018-08-10 Report/arne-report.Rmd
Modifies arnes report

## Descriptives
In normalised distance we can see similar starting point (not significant starting difference `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)

@@

There is difference in starting times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)). Probably due to learning issues with the treadmill. Threre is no significant difference in desktop and real (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))

@@

We can clearly see differences in the progression of the learnign, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling. I hope Mike and professor Ferer will help us with that.

But we can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.

@@

df_wide_errors$block14 <- df_wide_errors$block1-df_wide_errors$block4
df_wide_errors$block34 <- df_wide_errors$block3-df_wide_errors$block4
df_wide_distance$block14 <- df_wide_distance$block1-df_wide_distance$block4
df_wide_distance$block34 <- df_wide_distance$block3-df_wide_distance$block4

@@

block_t_test <- function(df, block1, block2, value){
  df <- df %>% filter(exp_block_id %in% c(block1,block2))
  mean.na <- function(data){return(mean(data, na.rm = T))}
  df_wide <- dcast(df, condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
  colnames(df_wide) <- c("condition","id", "block1","block2")
  df_wide %>% group_by(condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
}
```

@@

### Distance
We can see that all conditions improved from the first to forth block in both distance walked and errors in decisions. Paired t-tests show that both Treadmill and Real learning conditions led to improved performance from 1-4, but not Desktop. All conditions show improvemt in number of errors.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 1, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 1, 4,"errors")[,c(1,4)]
df14 <- merge(df_dist, df_error, by = "condition")
colnames(df14) <- c("condition", "p-value distance improvement", "p-value errors improvement")
kable(df14)
```

But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  nor errors (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`). This might be due to quite large variability in the sample.

@@

### Errors
```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

## Block improvement from block 3 to block 4

We can see that everybody (including Real to Real condition) gets a little bit worse from 3 to 4. But this "decrease of performance" is not significant when computed using paired t-tests except marginally worse in the desktop version. All conditions show imprvement in the number of errors in paired t-tests.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "condition")
colnames(df34) <- c("condition", "p-value distance improvement", "p-value errors improvement")
kable(df34)
```

Comparing the block improvement in different conditions using anovas, we see significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`)  and for errors (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

Runnign Tukey's post hoc on the distance shows marginally significant difference between Desktop from the other two conditions for distance.

```{r}
kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_distance), conf.level=0.95)$learning.condition)
```

And si
```{r}
kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition)
```

### Distance
```{r, echo=F}
ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
```

### Errors
```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
```

>>>>

## Descriptives
In normalised distance we can see that people in different conditions start fairly at the same level of performance (difference between learnign conditions is not significant `r apa_print(aov(min_norm_distance ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)

@@

There is difference in normalised times in all conditions in the first block (`r apa_print(aov(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1, ]))$statistic$learning_condition`)). Probably due to learning issues with the treadmill. Threre is no significant difference between desktop and real world learning in normalised times (`r apa_print(t.test(min_norm_time ~ learning.condition, conference_walk_all[conference_walk_all$exp_block_id == 1 & conference_walk_all$learning.condition %in% c("Desktop", "Real"), ]))$statistic`))

@@

We can clearly see differences in the progression of the learning slopes, but as you pointed out many times, as they are not linear, we could benefit from non linear modeling. I hope Mike and professor Ferer will help us with that.

But I can look at either pure changes between block 3-4 or 1-4 or those calculations that you proposed as a (block4-block3)/(block3+block4) measurements.

@@

df_wide_errors$block14 <- (df_wide_errors$block1-df_wide_errors$block4)/(df_wide_errors$block1+df_wide_errors$block4)
df_wide_errors$block34 <- (df_wide_errors$block3-df_wide_errors$block4)/(df_wide_errors$block3+df_wide_errors$block4)
df_wide_distance$block14 <- (df_wide_distance$block1-df_wide_distance$block4)/(df_wide_distance$block1+df_wide_distance$block4)
df_wide_distance$block34 <- (df_wide_distance$block3-df_wide_distance$block4)/(df_wide_distance$block3+df_wide_distance$block4)

@@

block_t_test <- function(df, block1, block2, value){
  df <- df %>% filter(exp_block_id %in% c(block1,block2))
  mean.na <- function(data){return(mean(data, na.rm = T))}
  df_wide <- dcast(df, learning.condition + id ~ exp_block_id, value.var = value, fun.aggregate = mean.na)
  colnames(df_wide) <- c("learning.condition","id", "block1","block2")
  df_wide %>% group_by(learning.condition) %>% do(tidy(t.test(.$block1, .$block2, paired = T)))
}

ez_prepare_block <- function(df, blocks){
  filtered_data <- df[df$exp_block_id %in% blocks, ]
  filtered_data <- filtered_data[complete.cases(filtered_data),]
  has_all <- filtered_data %>% group_by(id) %>% select(id, "exp_block_id") %>% unique() %>% count
  has_all <- has_all %>% filter(n == 2)
  filtered_data <- filtered_data %>% filter(id %in% has_all$id)
  filtered_data$exp_block_id <- factor(filtered_data$exp_block_id)
  return(filtered_data)
}

clean_walk_14 <- ez_prepare_block(conference_walk_all, c(1,4))
clean_walk_34 <- ez_prepare_block(conference_walk_all, c(3,4))
```

@@

### Distance
All conditions improved from the first to fourth block in both distance walked and errors in decisions. Paired t-tests show that both Treadmill and Real world learning conditions led to improved performance from 1-4, but Desktop did not. All conditions show significant improvement in number of errors from block 1 to 4.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 1, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 1, 4,"errors")[,c(1,4)]
df14 <- merge(df_dist, df_error, by = "learning.condition")
colnames(df14) <- c("learning condition", "p-value distance improvement", "p-value errors improvement")
kable(df14)
```

But anova shows no significant difference in the improvement between groups for distance (`r apa_print(aov(block14 ~ learning.condition, df_wide_distance))$statistic$learning_condition`). This might be due to quite large variability in the sample.

@@

### Errors
Unlike for distance, anova reveals significant difference in number of errors improvement from block 1 to block 4 between different learning conditions (`r apa_print(aov(block14 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

Runnign Tukey's post hoc on the distance shows that this is originating from the difference between Desktop and the Real world, with Treadmill not being significantly different from either of them.

```{r, echo=F}
kable(TukeyHSD(x=aov(block14 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
```

```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block14.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block14.diff-block14.se, ymax=block14.diff+block14.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement from block 1 to 4") + xlab("Learning condition")  + guides(fill=F)
```

## Block performance improvement from block 3 to block 4

Everybody's distance performance (including Real to Real condition) gets a little bit worse from 3 to 4. But this is not significant when computed using paired t-tests, although it is marginally worse in the desktop version. But all conditions show decrease in the number of errors from block 3 to 4.

```{r, echo=F}
df_dist <- block_t_test(conference_walk_all, 3, 4,"min_norm_distance")[,c(1,4)]
df_error <- block_t_test(conference_walk_all, 3, 4,"errors")[,c(1,4)]
df34 <- merge(df_dist, df_error, by = "learning.condition")
colnames(df34) <- c("learning.condition", "p-value distance improvement", "p-value errors improvement")
kable(df34)
```

### Distance

Comparing the block improvement in different conditions using anovas, we see marginally significant difference in the improvement between groups for distance (`r apa_print(aov(block34 ~ learning.condition, df_wide_distance))$statistic$learning_condition`).

```{r, echo=F}
ggplot(df_distance, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean normalised distance improvement from block 3 to 4") + xlab("Learning condition")  + guides(fill=F)
```

Within subject anova for distance improvement between 3 and 4 shows significant interaction effect and marginally significant effect of learning condition and block on distance.
```{r, warning = F, echo=F, message=F}
aov_block_dist_34 <- ezANOVA(
  data = clean_walk_34,
  dv = min_norm_distance, 
  wid = id,
  within = exp_block_id,
  between = learning.condition, 
  type = 3)
kable(aov_block_dist_34)

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = min_norm_distance, color = factor(learning.condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average normalised distance") + xlab("Experimental block") + 
  scale_color_manual(values = estimote_pal, "Learning condition")
```

### Errors

We can see significant differences between different conditions in error rate improvement from block 3 to 4 when running independent anova (`r apa_print(aov(block34 ~ learning.condition, df_wide_errors))$statistic$learning_condition`).

Post hoc tests show significant difference between Real and Desktop and marginally for Treadmill and Desktop.
```{r, echo=F}
kable(TukeyHSD(x=aov(block34 ~ learning.condition, df_wide_errors), conf.level=0.95)$learning.condition[, c(1,4)])
```

```{r, echo=F}
ggplot(df_errors, aes(x = learning.condition, y = block34.diff, fill = learning.condition)) + 
  geom_bar(stat = 'identity', position="dodge") + 
  geom_errorbar(aes(ymin=block34.diff-block34.se, ymax=block34.diff+block34.se), width=.2, position=position_dodge(.9)) +
  scale_fill_manual(values = estimote_pal) +
  ylab("Mean sum of errors improvement after switch")+ xlab("Learning condition") + guides(fill=F)
```

Running within subjects anova for block 3-4 number of errors shows significant effect of both condition and block, but unlike in distance improvement, their interaction doesn't seem to have an effect on error rate improvement.
```{r, warning = F, echo=F, message=F}
aov_block_err_34 <- ezANOVA(
  data = clean_walk_34,
  dv = errors, 
  wid = id,
  within = exp_block_id,
  between = condition,
  type = 3)
kable(aov_block_err_34)

ggplot(clean_walk_34, aes(x = as.numeric(exp_block_id), y = errors, color = factor(condition))) +
  stat_summary(fun.data=mean_cl_normal,position=position_dodge(width=0.1), width = 0.2, geom="errorbar") + 
  stat_summary(fun.y=mean,position=position_identity(), size = 1, geom="line") + 
  ylab("Average number of errors") + xlab("Experimental block") + 
  scale_color_manual(values = estimote_pal, "Learning condition")
```

## Summary
We see effect of learning condition (blocks 1-3) on performance in block 4. We also see

Participant's performance in distance from block 3-4 gets slightly worse after the switch, but their number of errors decrease. It can mean that they know which door are correct, but they are unsure in how to get to them, showing some level of memory transfer, but not necessarily spatial knowledge transfer.
